---
title: "The Beatles"
subtitle: "What are their top 5 popular songs?"
author: 
  - Hyunje Park, Charlie Zhang
thanks: "Code and data are available at: https://github.com/davidpxrk/beatles-top-5-songs"
date: today
date-format: long
abstract: "This paper examines which songs by the 1960s rock band The Beatles have the highest popularity in the United States, based on Spotify's popularity metric. It was discovered that the top 5 songs in popularity were Here Comes the Sun, Let It Be, Blackbird, Yesterday, In My Life."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false
#install.packages('ggplot2)
#install.packages('spotifyr')

library(spotifyr)
library(dplyr)
library(knitr)

# CHANGE THIS
Sys.setenv(SPOTIFY_CLIENT_ID = '')
Sys.setenv(SPOTIFY_CLIENT_SECRET = '')

access_token <- get_spotify_access_token()
```

\newpage

# Introduction

The Beatles, one of the most popular English rock bands of the 1960s, dominated the music industry. This paper analyzes which of their songs had the highest popularity ratings, using the Spotify API to gather data on their song performance. The data collection method is detailed in the Data section (@sec-data).

The paper is structured as follows: Data (@sec-data) introduces the dataset; Results (@sec-results) presents key findings; Discussion (@sec-discussion) explores observations and limitations; and Conclusion (@sec-conclusion) summarizes the analysis.

# Data {#sec-data}

This report uses the `Spotify API` to gather information on the popularity of `the Beatles`' songs. Each song in the Spotify API is represented in a `.json` format containing the variables `name` (name of the song) and `popularity` (a number from 0-100 indicating its popularity).

First, a Spotify Developer Account was created to get access to Spotify's `Web API`, and the client ID and secret ID were used to call Spotify's API. The `R Programming Language` [@citeR], along with the `spotifyr` [@citespotifyr], `dpylr` [@citedplyr] `knitr` [@citeknitr] packages was used to extract the top 5 songs of the band `The Beatles`, and make a presentable table.

# Results {#sec-results}

As shown in @fig-difference shows `the Beatles`' top 5 tracks. It was discovered that the top 5 songs in popularity were Here Comes the Sun, Let It Be, Blackbird, Yesterday, In My Life."

```{r}
#| label: fig-difference
#| fig-cap: "the Beatles' top 5 popular songs in the United States"
#| echo: false

beatles <- search_spotify('The Beatles', type = 'artist')

# Extract the artist ID (use the first result if it matches the name)
beatles_id <- beatles$id[1]


top_tracks <- get_artist_top_tracks(beatles_id, market = 'US')

# Extract and display the top 5 tracks
top_5 <- top_tracks %>%
  arrange(-popularity) %>%
  head(5) %>%
  select(name, popularity)
knitr::kable(top_5, format = "markdown", col.names = c("Song Name", "Popularity"))

```

\newpage

# Discussion {#sec-discussion}

The table @fig-difference showed the top 5 songs in popularity were Here Comes the Sun, Let It Be, Blackbird, Yesterday, In My Life. 

However, this is strict based off of `Spotify`'s popularity algorithm, that indicates a song's popularity based on recent play count, overall play count, user engagement and many other variables. This algorithm is kept hidden, hence it may be difficult to accurately reflect the popularity of the songs.


# Conclusion {#sec-conclusion}

In summary, this paper analyzed `the Beatles`' top 5 popular songs using the Spotify API. It was found that their top 5 songs in popularity were Here Comes the Sun, Let It Be, Blackbird, Yesterday, In My Life, respectively.


\newpage

# References
